---
title: Prerequisites 
---

Spaeth et al.'s (2024) Supreme Court Database (hereafter SCDB) provides scholars with a treaure trove of data on the U.S. Supreme Court. Political scientitst studying the Court frequently rely on the SCDB to find a host of information about cases and the justices. Acquring desired information from the SCDB often requires parsing through the database by executing lines of code. In this section, I have included a variety of lines of code that can be used on SCDB to acquire information, create different variables, and more.

Most code uses the Modern Database, which includes cases from the 1946-2023 terms. The code either use the Case Centered Citation data or the Justice Centered Citation data depending on the goal of the code.

To run the code, you need to download the latest version of both SCDB datasets which can be found at the SCDB's website [here](https://scdb.la.psu.edu/data). As of July 13, 2025, the latest data come from Version 2024 Release 01, released on October 1, 2024.

I use the .csv files of both dataset in R and the .dta files in Stata. In R, the Case Centered data object appears as `scdb_cc` and the Justice Centered data object appears as `scdb_jc`.

All data come from Spaeth et al. (2024). If you use this code to work with the SCDB, please properly cite Spaeth et al. (2024):

> Harold J. Spaeth, Lee Epstein, Michael J. Nelson, Andrew D. Martin, Jeffrey A. Segal, Theodore J. Ruger, and Sara C. Benesh. 2024. Supreme Court Database, Version 2024 Release 01. URL: [https://scdb.psu.edu](https://scdb.psu.edu)

# Loading in the data in R

Case Centered data
```{r eval = FALSE}
scdb_cc = read.csv("SCDB_2024_01_caseCentered_Citation.csv")
```
Justice Centered data 
```{r eval = FALSE}
scdb_jc = read.csv("SCDB_2024_01_justiceCentered_Citation.csv")
```

# Loading in the data in Stata
Case Centered data
```{r eval = FALSE}
use "SCDB_2024_01_caseCentered_Citation.dta", clear
```
Justice Centered data
```{r eval = FALSE}
use "SCDB_2024_01_justiceCentered_Citation.dta", clear
```


# References

> Harold J. Spaeth, Lee Epstein, Michael J. Nelson, Andrew D. Martin, Jeffrey A. Segal, Theodore J. Ruger, and Sara C. Benesh. 2024. Supreme Court Database, Version 2024 Release 01. URL: [https://scdb.psu.edu](https://scdb.psu.edu)
